The Evolution of Phenotypically Invariant Genetic Networks

### 2017-06-10 handnotes.tex
more notes on computing the gradient

Suppose that $\rho(t) \ge 0$ is a weighting function on $[0,\infty)$
so that fitness is a function of $L^2(\rho)$ distance of the impulse response from optimal:
\begin{equation}
    \begin{aligned}
        D(A) 
        &:= 
        \int_0^\infty \rho(t) \left| C e^{At} B - C e^{A_0 t} B \right|^2 dt \\
        &= 
        \int_0^\infty \rho(t) \left| C \left( e^{At} - e^{A_0 t} \right) B \right|^2 dt \\
        &= 
        \int_0^\infty \rho(t) C \left( e^{At} - e^{A_0 t} \right) B B^T \left( e^{At} - e^{A_0 t} \right)^T C^T dt
    \end{aligned}
\end{equation}
How does this change with $A$?
Since
\begin{equation}
  \begin{aligned}
      \frac{d}{du} e^{(A+uZ)t} \vert_{u=0}
      &=
      \int_0^t e^{As} Z e^{A(t-s)} ds, 
  \end{aligned}
\end{equation}
we have that
\begin{equation}
  \begin{aligned}
      \frac{d}{du} D(A+uZ)\vert_{u=0}
      &=
        2 \int_0^\infty \rho(t) C \left( \int_0^t e^{As} Z e^{A(t-s)} ds \right) B B^T \left( e^{At} - e^{A_0 t} \right)^T C^T dt \\
      &=
        2 \int_0^\infty \int_0^t \rho(t) C e^{As} Z e^{A(t-s)} B B^T \left( e^{At} - e^{A_0 t} \right)^T C^T ds dt .
  \end{aligned}
\end{equation}

>>>>

Suppose that $\rho(t) \ge 0$ is a weighting function on $[0,\infty)$
so that fitness is a function of $L^2(\rho)$ distance of the impulse response from optimal.
With $A_0$ a representative of the optimal set:
\begin{equation}
    \begin{aligned}
        D(A) 
        &:= 
        \int_0^\infty \rho(t) \left| h_A(t) - h_{A_0}(t) \right|^2 dt \\
        &:= 
        \int_0^\infty \rho(t) \left| C e^{At} B - C e^{A_0 t} B \right|^2 dt \\
        &= 
        \int_0^\infty \rho(t) \left| C \left( e^{At} - e^{A_0 t} \right) B \right|^2 dt \\
        &= 
        \int_0^\infty \rho(t) C \left( e^{At} - e^{A_0 t} \right) B B^T \left( e^{At} - e^{A_0 t} \right)^T C^T dt
    \end{aligned}
\end{equation}
How does this change with $A$?
Since
\begin{equation}
  \begin{aligned}
      \frac{d}{du} e^{(A+uZ)t} \vert_{u=0}
      &=
      \int_0^t e^{As} Z e^{A(t-s)} ds, 
  \end{aligned}
\end{equation}
we have that
\begin{equation}
  \begin{aligned}
      \frac{d}{du} D(A+uZ)\vert_{u=0}
      &=
        2 \int_0^\infty \rho(t) C \left( \int_0^t e^{As} Z e^{A(t-s)} ds \right) B B^T \left( e^{At} - e^{A_0 t} \right)^T C^T dt \\
      &=
        2 \int_0^\infty \rho(t) C \left( \int_0^t e^{As} Z e^{A(t-s)} ds \right) B \left( h_A(t) - h_{A_0}(t) \right)^T dt 
  \end{aligned}
\end{equation}

We could use this to compute the gradient of $D$.
To do so, it'd be nice to have a way to compute the inner integral above.
Suppose that we can diagonalize $A = U \Lambda U^{-1}$.
Then
\begin{equation} \label{eqn:exp_deriv}
  \begin{aligned}
      \int_0^t e^{As} Z e^{A(t-s)} ds 
      &=
      \int_0^t U e^{\Lambda s} U^{-1} Z U e^{\Lambda (t-s)} U^{-1} ds 
  \end{aligned}
\end{equation}
Now, notice that
\begin{equation}
  \begin{aligned}
      \int_0^t e^{s \lambda_i} e^{(t-s) \lambda_j} ds
      &=
      \frac{ e^{t \lambda_i} - e^{t \lambda_j} }{ \lambda_i - \lambda_j } .
  \end{aligned}
\end{equation}
Therefore, 
defining
\begin{equation}
    X_{ij}(t) = \left( U^{-1} Z U \right)_{ij}
      \frac{ e^{t \lambda_i} - e^{t \lambda_j} }{ \lambda_i - \lambda_j } 
\end{equation}
moving the $U$ and $U^{-1}$ outside the integral and integrating we get that
\begin{equation}
  \begin{aligned}
      \int_0^t e^{As} Z e^{A(t-s)} ds 
      &=
      U X(t) U^{-1} .
  \end{aligned}
\end{equation}
